Turbulence and Particle Acceleration in Giant Radio Haloes: the Origin of Seed Electrons

### 2016-06-23 paper.tex
more edits; open points in section 4.3

@@

Given a seed population of CRs, we adopt essentially the same set of
plasma physics assumptions as the reacceleration model for RHs
\citep{brunetti07,brunetti11}. We solve the isotropic, gyro-phase
averaged Fokker-Planck equation (via a Crank-Nicholson scheme) for the
time evolution of the CRe distribution in the Lagrangian frame
\citep{brunetti07,brunetti11}:
\begin{eqnarray}
{{d f_{\rmn{e}}(p,t)}\over{d t}} &\!=&
\frac{\partial}{\partial p}
\left\{
f_{\rmn{e}}(p,t)\left[
\left|\frac{d p}{d t}\right|_{\rm Coul} 
+ \frac{p}{3}\left(\bnabla\bcdot \bvel\right)
+ \left|\frac{d p}{d t}\right|_{\rm rad}\right.\right.
\nonumber\\
&-& \left.\left.{1\over{p^2}}{{\partial }\over{\partial p}}\left(p^2 D_{pp}\right) 
\right]\right\} - \left(\bnabla\bcdot \bvel\right) f_{\rmn{e}}(p,t)
\nonumber\\
&+& {{\partial^2 }\over{\partial p^2}}
\left[
D_{pp} f_{\rmn{e}}(p,t) \right]+ Q_{\rmn{e}}\left[p,t;f_{\rmn{p}}(p,t)\right]   \, .
\label{elettroni}
\end{eqnarray}
Here $f_{\rmn{e}}$ is the one-dimensional distribution in position $x$
(suppressed for clarity), momentum $p=P/(m_{\rmn{e}} c)$ and time $t$ (which is
normalised such that the number density is given by
$n_{\rmn{CRe}}(t)=\int d p f_{\rmn{e}}(p,t)$), $d/d t=\partial/\partial
t+\bvel\bcdot\bnabla$ is the Lagrangian derivative, $\bvel$ is the gas
velocity, $|d p/d t|$ represents Coulomb
\citep[Coul,][]{1972Phy....60..145G} and radiative
\citep[rad,][]{1979rpa..book.....R} losses, respectively,
\begin{eqnarray}
   \left|\frac{d p}{d t}\right|_{\rm Coul} &=& \frac{3\,\sigma_\rmn{T}\,n_{\rmn{e}}\,c}{2\, \beta_{\rmn{e}}^2}
  \left[\ln\left(\frac{m_{\rmn{e}} c^2 \beta_{\rmn{e}} \sqrt{\gamma-1}}{\hbar\,\omega_\rmn{plasma}}\right)\right.\nonumber \\
    &-&\ln(2)\left(\frac{\beta_{\rmn{e}}^2}{2}+\frac{1}{\gamma}\right)+\frac{1}{2}+
    \left.\left(\frac{\gamma-1}{4\gamma}\right)^2\right]\,, \\
  \left|\frac{d p}{d t}\right|_{\rm rad~} &=& \frac{4}{3}\frac{\sigma_\rmn{T}}{m_{\rmn{e}}c}\,
  \frac{p^2}{\beta_{\rmn{e}}}\,\left[1+\left(\frac{B}{B_\rmn{CMB}}\right)^2\right]\,\eps_{\rmn{CMB}}\,.
\end{eqnarray}
Here $\beta_{\rmn{e}} = p/\sqrt{1+p^2}$ is the dimensionless velocity of CRs,
$\gamma=\sqrt{1+p^2}$ is the Lorentz factor of CRs,
$\omega_\rmn{plasma} = \sqrt{4\upi e^2 n_{\rmn{e}} / m_{\rmn{e}}}$ is the plasma
frequency, $n_e$ is the number density of free electrons, and
$\sigma_\rmn{T}= 8\upi e^4/3(m_{\rmn{e}} c^2)^2$ is the Thomson cross
section. The {\it rms} magnetic field strength is denoted by $B$ and
the equivalent field strength of the cosmic-microwave background is
given by $B_\rmn{CMB} = 3.24 (1 + z)^2\umu\rmn{G}$, where $z$ denotes
the redshift. In the peripheral cluster regions, where $B \ll
B_\rmn{CMB}$, the CRes loose virtually all their energy by means of
inverse Compton emission. $D_{pp}$ is the momentum space diffusion
coefficient (see \S\ref{sec:reacc}), and
$Q_{\rmn{e}}$ denotes the injection rate of primary and secondary
electrons in the ICM (see section~\ref{sec:cosmo_sim}). The first term
containing the expression $\bnabla\bcdot \bvel$ represents Fermi-I
acceleration and the second term of this form describes adiabatic
gains and losses.

@@

The time evolution of the spectral energy distribution of CRps,
$f_{\rmn{p}}$, is similarly given by:
\begin{eqnarray}
\lefteqn{
  {{d f_{\rmn{p}}(p,t)}\over{d t}} =
  {{\partial }\over{\partial p}}
  \left\{
  f_{\rmn{p}}(p,t)\left[ \left|{{d p}\over{d t}}\right|_{\rm Coul}
    + \frac{p}{3}\bnabla\bcdot \left(\bvel+\bvel_{\rmn{st}}\right)\right.\right.}
\nonumber\\
&&-\left.\left. {1\over{p^2}}{{\partial }\over{\partial p}}\left(p^2 D_{pp}\right)
\right]\right\} - f_{\rmn{p}}(p,t) \bnabla\bcdot \left(\bvel+\bvel_{\rmn{st}}\right) 
- \bvel_{\rmn{st}}\bcdot\bnabla f_{\rmn{p}}(p,t)
\nonumber\\
&&+\,\,\, {{\partial^2 }\over{\partial p^2}}
\left[ D_{pp} f_{\rmn{p}}(p,t) \right] - {{f_{\rmn{p}}(p,t)}\over{\tau_{\rm had}(p)}}
+ Q_{\rmn{p}}(p,t)\, ,
\label{eq:FP_p}
\end{eqnarray}
where $\bvel_{\rmn{st}}=-\varv_{\rmn{A}}\bnabla f_{\rmn{p}}/|\bnabla
f_{\rmn{p}}|$ is the streaming velocity in the isotropic transport
approximation, $\varv_{\rmn{A}}$ is the Alfv\'en speed,
and $Q_{\rmn{p}}(p,t)$ denotes the injection rate of shock accelerated
CRps as a function of momentum $p=P/(m_{\rmn{p}} c)$ and time $t$ (see
section~\ref{sec:cosmo_sim}). The timescale of
hadronic losses that produce pions via CRp collisions with thermal
protons of the ICM is given by 
\begin{eqnarray}
  \tau_{\rm had} = \left[c\,n_\rmn{th}\,\sigma^{+/-,0}(p)\right]^{-1}\,,
\end{eqnarray}
where we use the cross-section, $\sigma^{+/-,0}(p)$, given by the
fitting formula in \citet{1986ApJ...307...47D}.

@@

All of the physics of turbulent reacceleration is effectively encapsulated in the diffusion coefficient \citep{brunetti07}, which can be rewritten as \citep{miniati15}:
\begin{equation}
D_{pp}(p) = \frac{ p^{2} \upi I_{\theta}(c_{s}/c)}{8 c} \langle k \rangle_{\mathcal{W}} \langle (\delta \varv_{\rm c})^{2} \rangle
\label{eqn:diffusion}
\end{equation}
where $I_{\theta}$ averages interaction rates over the CR pitch angle $\theta$; $I_{\theta} \approx 5$ for ICM conditions, and
\begin{equation}
\langle k \rangle_{\mathcal{W}} = \frac{1}{\langle (\delta \varv_{\rm c})^{2} \rangle} \int_{k_L}^{k_{\rm cut}} d k \, k \, \mathcal{W}(k) \approx \frac{s-1}{2-s} k_L \left( \frac{k_{\rm cut}}{k_L} \right)^{2-s} 
\label{eqn:k_W} 
\end{equation}
is an energy-averaged wavenumber, $k_L, k_{\rm cut}$ are the wavenumbers associated with the outer scale $L$ and the cutoff scale respectively, and we have assumed a total energy spectrum (composed of both kinetic and potential energy, where the two are assumed to be in equipartition, \citealt{sarkar11}): 
\begin{equation}
\mathcal{W}(k) = \frac{(s-1) \langle (\delta \varv_{\rm c})^{2} \rangle}{k_L} \left( \frac{k}{k_L} \right)^{-s} 
\label{eqn:Wk}
\end{equation}
which defines the normalisation $\langle (\delta \varv_{\rm c})^{2} \rangle$ (the subscript 'c' emphasizes that we specialise to compressive modes). Intuitively, we can understand the form of the the diffusion coefficient from the fact that for second-order Fermi acceleration, $\dot{p} \sim \Delta p/\tau \sim p\,\langle k c \rangle_{\mathcal{W}} (\delta \varv_{\rm c}/c)^{2}$, where $\tau^{-1} \sim \langle k c \rangle_{\mathcal{W}}$ is the energy averaged wave-particle interaction rate, and $\Delta p \sim p (\delta \varv_{\rm c}/c)^{2}$ is the typical momentum change during wave-particle scattering. Thus: 
\begin{equation}
D_{pp} \sim p \dot{p} \sim p^{2} \langle k c \rangle_{\mathcal{W}} \left( \frac{\delta \varv_{\rm c}}{c} \right)^{2}. 
\end{equation}

Equations \ref{eqn:diffusion} and \ref{eqn:Wk} make the important aspects of turbulence clear: the energy in compressive modes $\langle (\delta \varv_{\rm c})^{2} \rangle$, the inner and outer scale ($k_{\rm cut}$ and $k_L$), and the slope of the energy spectrum $s$. All of these can vary spatially and temporally.

We adopt a Kraichnan spectrum ($s=3/2$) for the fast modes, as seen in simulations \citep{cho03}. This can also be written as: 
\begin{equation}
  \label{eq:Wk}
  \mathcal{W}(k) =
\sqrt{2/7\,I_L\,\rho\,\langle \Vph \rangle}\,k^{-3/2},
\end{equation}
where $I_L$ is the volumetric injection rate of turbulence at scale $L$ (which is assumed to be constant), $\varv_{\rm ph}\approx c_{\rm s}$ is the phase speed of waves. It is important to realise that this only applies at small scales, where turbulence is magneto-hydrodynamical. At large scales, turbulent motions are subsonic (${\mathcal M}_{\rm s} \sim 0.2-0.5$) but super-Alfv{\'e}nic (${\mathcal M}_{\rm A} \sim 5$). Thus, except in the case where motions are transsonic and weak shocks become important, motions are fundamentally hydrodynamic and turbulence follows a Kolmogorov ($s=5/3$) spectrum. However, while the turbulent energy density $\eps_{\rm turb} \propto l^{2/3}$ decreases at small scales, the magnetic energy density is scale-independent. Thus, at some scale $l_{\rm A} \sim l_L {\mathcal M}_{\rm A}^{-3}$, where $\eps_{\rm turb} \sim \eps_{B}$, the magnetic field becomes dynamically important, and turbulence transitions to the MHD regime. The cutoff scale in equation \ref{eqn:Wk} can be computed by setting the turbulent cascade rate for fast modes to the transit time damping rate on thermal electrons. This yields \citep{brunetti07,miniati15}:
\begin{equation}
k_{\rm cut} \approx A k_L \frac{\langle \varv_L^{2}\rangle^{2}}{c_{\rm s}^{4}}
%k_{\rm c} \approx A k_{\rm outer} \frac{\langle \varv_{\rm outer}^{2}\rangle^{2}}{c_{\rm s}^{4}}
\label{eqn:k_c} 
\end{equation}
where $A \approx 11 000$. This gives $2\upi/k_\rmn{cut}\sim 0.1-1$~kpc in the ICM. This constitutes an
effective mean free path for CRs, unless plasma instabilities can
mediate interactions between turbulence and particles on smaller
scales \citep{brunetti11}, a possibility we discuss in \S\ref{sect:self-limiting}. Another possibility is that compressible modes dissipate in weak shocks, resulting in Burgers' turbulence $s=2$ \citep{kowal10,porter15,miniati15}. If Burgers turbulence dominates, then particle acceleration rates are too slow in the face of cooling processes to explain radio haloes \citep{miniati15}, and an alternative model for radio haloes is required.

@@

The important effects are best summarised in terms of the acceleration time: 
\begin{equation}
\tau_{\rm D}= \frac{p^{2}}{4 D_{pp}}
\end{equation}
which is momentum independent. This should be compared against the lifetime of turbulence, $\tau_{\rm cl}$, and the cooling time $\tau_{\rm cool}$. In Table~\ref{tab:timescales}, we show both the thermal quantities and the
timescales for CR cooling and (re)acceleration for three different
spatial regions of the RH. The densities \citep{1992A&A...259L..31B} and temperatures \citep{2009ApJ...696.1886B,2001A&A...365L..67A} are derived from X-ray observations. To calculate synchrotron cooling times, we use $B$-fields derived from Faraday rotation measurements \citep{bonafede10}. To calculate the acceleration time, we need to assume an outer scale. We assume an injection scale $k_L=2\upi/\lambda_L$, where $\lambda_L = 100$ kpc, which were assumptions adopted in previous work \citep{2006MNRAS.366.1437S,brunetti07, brunetti11}. This length scale corresponds to an eddy turnover time on the outer scale of $\varv_L k_L^{-1} \sim 1.2$~Gyr if $\varv_L \sim 500 {\rm km \, s^{-1}}$, as is characteristic of a merger. Note that hydrodynamical simulations of clusters have sometimes found larger $\lambda_L$, in some cases comparable to the size of the cluster \citep{vazza11,miniati15}. This choice is degenerate with $X_{\rm tu}$; in \S\ref{sect:self-limiting} we argue that $k_{\rm A}$ is a more appropriate choice, but also find that when the decay time is appropriately scaled, we are relatively insensitive to the choice of $k_L$. We present $\tau_{\rm D}$ for 3 different models, which we discuss in the next section. The reacceleration timescale
$\tau_{\rm D}$ is similar between our three models, where the
difference comes from turbulent profile parametrized with
$\alpha_\rmn{tu}$. This implies that even small differences in the
turbulent profile could impact the seed CRs significantly. Finally, we adopt an duration of acceleration $\tau_{\rm cl} = 650$ Myr, in line with previous assumptions \citep{brunetti07}, roughly corresponding to the turbulent decay time. 
\begin{table}
  \caption{Thermal quantities and timescales for different spatial
    regions in a Coma like cluster.}
\begin{tabular}{l c  c c}
\hline
\hline
& & spatial regions & \\
 & $0.1\,\RH^{(2)}$ & $0.3\,\RH^{(2)}$ & $\RH^{(2)}$   \\
\hline
thermal quantities$^{(1)}$ & & & \\
$\rho$ [$10^{-27}$ g cm$^{-3}$] & 3.0 & 1.6 & 0.15 \\
$T$ [$10^{8}$ K] & 1.4 & 1.0 & 0.58 \\
\hline
timescales$^{(1)}$ & & & \\
$\tau_\rmn{D}$(\Mprimary)$^{(3)}$ [Gyr] & 0.45 & 0.44 & 0.39 \\
$\tau_\rmn{D}$(\Mstream)$^{(3)}$  [Gyr] & 0.50 & 0.47 & 0.34 \\
$\tau_\rmn{D}$(\Mflatturb)$^{(3)}$  [Gyr] & 0.69 & 0.56 & 0.27 \\
$\tau_\rmn{IC/sync}(P=10^4\,m_{\rmn{e}}c)^{(4)}$ [Gyr] & 0.11 & 0.15 & 0.22 \\
$\tau_\rmn{had}(P=100\,m_{\rmn{p}}c)^{(5)}$ [Gyr] & 2.4 & 4.5 & 47 \\
$\tau_\rmn{Coul}(P=\,m_{\rmn{e}}c)^{(6)}$  [Gyr] & 0.0092 & 0.017 & 0.17 \\
\hline
\end{tabular}
\begin{quote}
 Notes: \\ 
 (1) Median quantities from our simulated post-merging cluster g72a derived 
 during last 300~Myrs in time. \\
 (2) Radius of the giant radio halo in Coma where $\RH\approx0.6\,R_{200}$.\\
 (3) Fermi-II reacceleration for both electrons and protons at all energies.\\
 (4) Inverse Compton and synchrotron cooling for electrons.\\
 (5) Catastrophic losses for protons.\\
 (6) Coulomb cooling for electrons (protons factor $m_{\rmn{e}}/m_{\rmn{p}}$ smaller).\\
 \label{tab:timescales}
  \end{quote}
\end{table}

>>>>

\usepackage[normalem]{ulem}
\usepackage[usenames]{xcolor}
\newcommand\cp[1]{{\color{red} #1}}
\newcommand\po[1]{{\color{blue} #1}}

@@

Given a seed population of CRs, we adopt essentially the same set of
plasma physics assumptions as the reacceleration model for RHs
\citep{brunetti07,brunetti11}. We solve the isotropic, gyro-phase
averaged Fokker-Planck equation (via a Crank-Nicholson scheme) for the
time evolution of the CRe distribution in the Lagrangian frame
\citep{brunetti07,brunetti11}:
\begin{eqnarray}
{{d f_{\rmn{e}}(p,t)}\over{d t}} &\!=&
\frac{\partial}{\partial p}
\left\{
f_{\rmn{e}}(p,t)\left[
\left|\frac{d p}{d t}\right|_{\rm Coul} 
+ \frac{p}{3}\left(\bnabla\bcdot \bvel\right)
+ \left|\frac{d p}{d t}\right|_{\rm rad}\right.\right.
\nonumber\\
&-& \left.\left.{1\over{p^2}}{{\partial }\over{\partial p}}\left(p^2 D_{pp}\right) 
\right]\right\} - \left(\bnabla\bcdot \bvel\right) f_{\rmn{e}}(p,t)
\nonumber\\
&+& {{\partial^2 }\over{\partial p^2}}
\left[
D_{pp} f_{\rmn{e}}(p,t) \right]+ Q_{\rmn{e}}\left[p,t;f_{\rmn{p}}(p,t)\right]   \, .
\label{elettroni}
\end{eqnarray}
Here $f_{\rmn{e}}$ is the one-dimensional distribution in position $x$
(suppressed for clarity), momentum \cp{$p=P/(m_{\rmn{e}} c)$} and time $t$
(which is normalised such that the number density is given by
$n_{\rmn{CRe}}(t)=\int d p f_{\rmn{e}}(p,t)$), $d/d t=\partial/\partial
t+\bvel\bcdot\bnabla$ is the Lagrangian derivative, $\bvel$ is the gas velocity,
$|d p/d t|$ represents Coulomb \citep[Coul,][]{1972Phy....60..145G} and
radiative \citep[rad,][]{1979rpa..book.....R} losses, respectively,
\begin{eqnarray}
   \left|\frac{d p}{d t}\right|_{\rm Coul} &=& \frac{3\,\sigma_\rmn{T}\,n_{\rmn{e}}\,c}{2\, \beta_{\rmn{e}}^2}
  \left[\ln\left(\frac{m_{\rmn{e}} c^2 \beta_{\rmn{e}} \sqrt{\gamma-1}}{\hbar\,\omega_\rmn{plasma}}\right)\right.\nonumber \\
    &-&\ln(2)\left(\frac{\beta_{\rmn{e}}^2}{2}+\frac{1}{\gamma}\right)+\frac{1}{2}+
    \left.\left(\frac{\gamma-1}{4\gamma}\right)^2\right]\,, \\
  \left|\frac{d p}{d t}\right|_{\rm rad~} &=& \frac{4}{3}\frac{\sigma_\rmn{T}}{m_{\rmn{e}}c}\,
  \frac{p^2}{\beta_{\rmn{e}}}\,\left[1+\left(\frac{B}{B_\rmn{CMB}}\right)^2\right]\,\cp{\eps_{\rmn{CMB}}}\,.
\end{eqnarray}
Here $\beta_{\rmn{e}} = p/\sqrt{1+p^2}$ is the dimensionless velocity of CRs,
$\gamma=\sqrt{1+p^2}$ is the Lorentz factor of CRs,
$\omega_\rmn{plasma} = \sqrt{4\upi e^2 n_{\rmn{e}} / m_{\rmn{e}}}$ is the plasma
frequency, $n_e$ is the number density of free electrons, and
$\sigma_\rmn{T}= 8\upi e^4/3(m_{\rmn{e}} c^2)^2$ is the Thomson cross
section. The {\it rms} magnetic field strength is denoted by $B$ and
the equivalent field strength of the cosmic-microwave background is
given by $B_\rmn{CMB} = 3.24 (1 + z)^2\umu\rmn{G}$, where $z$ denotes
the redshift. In the peripheral cluster regions, where $B \ll
B_\rmn{CMB}$, the CRes loose virtually all their energy by means of
inverse Compton emission. $D_{pp}$ is the momentum space diffusion
coefficient (see \S\ref{sec:reacc}), and
$Q_{\rmn{e}}$ denotes the injection rate of primary and secondary
electrons in the ICM (see section~\ref{sec:cosmo_sim}). The first term
containing the expression $\bnabla\bcdot \bvel$ represents Fermi-I
acceleration and the second term of this form describes adiabatic
gains and losses.

@@

The time evolution of the spectral energy distribution of CRps,
$f_{\rmn{p}}$, is similarly given by:
\begin{eqnarray}
\lefteqn{
  {{d f_{\rmn{p}}(p,t)}\over{d t}} =
  {{\partial }\over{\partial p}}
  \left\{
  f_{\rmn{p}}(p,t)\left[ \left|{{d p}\over{d t}}\right|_{\rm Coul}
    + \frac{p}{3}\bnabla\bcdot \left(\bvel+\bvel_{\rmn{st}}\right)\right.\right.}
\nonumber\\
&&-\left.\left. {1\over{p^2}}{{\partial }\over{\partial p}}\left(p^2 D_{pp}\right)
\right]\right\} - f_{\rmn{p}}(p,t) \bnabla\bcdot \left(\bvel+\bvel_{\rmn{st}}\right) 
- \bvel_{\rmn{st}}\bcdot\bnabla f_{\rmn{p}}(p,t)
\nonumber\\
&&+\,\,\, {{\partial^2 }\over{\partial p^2}}
\left[ D_{pp} f_{\rmn{p}}(p,t) \right] - {{f_{\rmn{p}}(p,t)}\over{\tau_{\rm had}(p)}}
+ Q_{\rmn{p}}(p,t)\, ,
\label{eq:FP_p}
\end{eqnarray}
where $\bvel_{\rmn{st}}=-\varv_{\rmn{A}}\bnabla f_{\rmn{p}}/|\bnabla
f_{\rmn{p}}|$ is the streaming velocity in the isotropic transport
approximation, $\varv_{\rmn{A}}$ is the Alfv\'en speed,
and $Q_{\rmn{p}}(p,t)$ denotes the injection rate of shock accelerated
CRps as a function of momentum \cp{$p=P/(m_{\rmn{p}} c)$} and time $t$ (see
section~\ref{sec:cosmo_sim}). The timescale of
hadronic losses that produce pions via CRp collisions with thermal
protons of the ICM is given by 
\begin{eqnarray}
  \tau_{\rm had} = \left[c\,n_\rmn{th}\,\sigma^{+/-,0}(p)\right]^{-1}\,,
\end{eqnarray}
where we use the cross-section, $\sigma^{+/-,0}(p)$, given by the
fitting formula in \citet{1986ApJ...307...47D}.

@@

All of the physics of turbulent reacceleration is effectively encapsulated in the diffusion coefficient \citep{brunetti07}, which can be rewritten as \citep{miniati15}:
\begin{equation}
D_{pp}(p) = \frac{ p^{2} \upi I_{\theta}(c_{s}/c)}{8 c} \langle k \rangle_{\mathcal{W}} \langle (\delta \varv_{\rm c})^{2} \rangle
\label{eqn:diffusion}
\end{equation}
where $I_{\theta}$ averages interaction rates over the CR pitch angle $\theta$; $I_{\theta} \approx 5$ for ICM conditions, and
\begin{equation}
\langle k \rangle_{\mathcal{W}} = \frac{1}{\langle (\delta \varv_{\rm c})^{2} \rangle} \int_{k_L}^{k_{\rm cut}} d k \, k \, \mathcal{W}(k) \approx \frac{s-1}{2-s} k_L \left( \frac{k_{\rm cut}}{k_L} \right)^{2-s} 
\label{eqn:k_W} 
\end{equation}
is an energy-averaged wavenumber, $k_L, k_{\rm cut}$ are the wavenumbers associated with the outer scale $L$ and the cutoff scale respectively, and we have assumed a total energy spectrum (composed of both kinetic and potential energy, where the two are assumed to be in equipartition, \citealt{sarkar11}): 
\begin{equation}
\mathcal{W}(k) = \frac{(s-1) \langle (\delta \varv_{\rm c})^{2} \rangle}{k_L} \left( \frac{k}{k_L} \right)^{-s} 
\label{eqn:Wk}
\end{equation}
which defines the normalisation $\langle (\delta \varv_{\rm c})^{2} \rangle$ (the subscript 'c' emphasizes that we specialise to compressive modes). Intuitively, we can understand the form of the the diffusion coefficient from the fact that for second-order Fermi acceleration, $\dot{p} \sim \Delta p/\tau \sim p\,\langle k c \rangle_{\mathcal{W}} (\delta \varv_{\rm c}/c)^{2}$, where $\tau^{-1} \sim \langle k c \rangle_{\mathcal{W}}$ is the energy averaged wave-particle interaction rate, and $\Delta p \sim p (\delta \varv_{\rm c}/c)^{2}$ is the typical momentum change during wave-particle scattering. Thus: 
\begin{equation}
  \label{eqn:Dpp}
D_{pp} \sim p \dot{p} \sim p^{2} \langle k c \rangle_{\mathcal{W}} \left( \frac{\delta \varv_{\rm c}}{c} \right)^{2}. 
\end{equation}

Equations (\ref{eqn:diffusion}) and (\ref{eqn:Wk}) make the important aspects of turbulence clear: the energy in compressive modes $\langle (\delta \varv_{\rm c})^{2} \rangle$, the inner and outer scale ($k_{\rm cut}$ and $k_L$), and the slope of the energy spectrum $s$. All of these can vary spatially and temporally.

We adopt a Kraichnan spectrum ($s=3/2$) for the fast modes, as seen in simulations \citep{cho03}. This can also be written as: 
\begin{equation}
  \label{eq:Wk}
  \mathcal{W}(k) =
\sqrt{2/7\,I_L\,\rho\,\langle \Vph \rangle}\,k^{-3/2},
\end{equation}
where $I_L$ is the volumetric injection rate of turbulence at scale $L$ (which is assumed to be constant), $\varv_{\rm ph}\approx c_{\rm s}$ is the phase speed of waves. \cp{Since the Kraichnan spectrum is critical in what follows, it is worthwhile taking a moment to recount the origin of the spectral slope $s$ and the cascade rate. In the magnetically dominated regime, two counterpropagating wave packets of scale $l$ interact on an Alfv{\'e}n time $\tau_{\rm A}=l/\varv_{\rm A}$, rather than the eddy turnover time $\tau_{\rm edd} = l/\varv_{\rm l}$. 
Since $\tau_{\rm A} \ll \tau_{\rm edd}$, each interaction results in a small velocity change $\delta \varv_{\rm l} \sim \varv_{\rm l} (\tau_{\rm A}/\tau_{\rm edd})$. If these changes behave like a random walk, the cascade time $\tau_{\rm l}$ it takes for an eddy to become non-linear ($\Delta \varv \sim \varv_{\rm l})$ and cascades to smaller scales occurs with a characteristic velocity $\varv_{\rm l} \sim \delta \varv_{\rm l} (\tau_{\rm l}/\tau_{\rm A})^{1/2} \sim \varv_{\rm l} \tau_{\rm A}/\tau_{\rm edd} (\tau_{\rm l}/\tau_{\rm A})^{1/2}$, which we can solve to obtain the cascade time: 
\begin{equation}
  \label{eqn:tau_l}
\tau_{\rm l} \sim \left(\frac{\tau_{\rm edd}}{\tau_{\rm A}}\right)^{2} \tau_{\rm A} \sim \frac{l \varv_{\rm A}}{\varv_{\rm l}^{2}} 
\end{equation}
where we have implicitly assumed isotropy. This implies a dissipation rate $\epsilon \sim \varv_{\rm l}^{2}/\tau_{\rm l} \sim \varv_{\rm l}^{4}/(l \varv_{\rm A})$, or $\varv_{\rm l} \sim (\epsilon l \varv_{A})^{1/4}$. Thus, since $k E(k)\sim \varv_{\rm l}^{2}$, this gives a kinetic energy spectrum: 
\begin{equation}
E(k) \sim (\epsilon \varv_{\rm A})^{1/2} k^{-3/2}.
\end{equation}}

It is important to realise that this Kraichnan scaling only applies at small scales, where turbulence is magneto-hydrodynamical. At large scales, turbulent motions are subsonic (${\mathcal M}_{\rm s} \sim 0.2-0.5$) but super-Alfv{\'e}nic (${\mathcal M}_{\rm A} \sim 5$). Thus, except in the case where motions are transsonic and weak shocks become important, motions are fundamentally hydrodynamic and turbulence follows a Kolmogorov ($s=5/3$) spectrum. However, while the turbulent energy density $\eps_{\rm turb} \propto l^{2/3}$ decreases at small scales, the magnetic energy density is scale-independent. Thus, at some scale $l_{\rm A} \sim l_L {\mathcal M}_{\rm A}^{-3}$, where $\eps_{\rm turb} \sim \eps_{B}$, the magnetic field becomes dynamically important, and turbulence transitions to the MHD regime. The cutoff scale in equation (\ref{eqn:Wk}) can be computed by setting the turbulent cascade rate for fast modes to the transit time damping rate on thermal electrons. This yields \citep{brunetti07,miniati15}:
\begin{equation}
k_{\rm cut} \approx A k_L \frac{\langle (\delta \varv_{\rmn{c}})^{2}\rangle^{2}}{c_{\rm s}^{4}}
\label{eqn:k_c} 
\end{equation}
where $A \approx 11 000$. This gives $2\upi/k_\rmn{cut}\sim 0.1-1$~kpc in the ICM. This constitutes an
effective mean free path for CRs, unless plasma instabilities can
mediate interactions between turbulence and particles on smaller
scales \citep{brunetti11}, a possibility we discuss in \S\ref{sect:self-limiting}. Another possibility is that compressible modes dissipate in weak shocks, resulting in Burgers' turbulence $s=2$ \citep{kowal10,porter15,miniati15}. If Burgers turbulence dominates, then particle acceleration rates are too slow in the face of cooling processes to explain radio haloes \citep{miniati15}, and an alternative model for radio haloes is required.

@@

The important effects are best summarised in terms of the acceleration \cp{rate
  which is goverend by advection in momentum space:
\begin{equation}
  \tau_{\rm D}^{-1}\equiv \Gamma_{\rm D}\equiv
  \frac{\dot{p}}{p}=p^{-3}\frac{\partial}{\partial p}\left(p^2D_{pp}\right)
  =\frac{4 D_{pp}}{p^{2}}.
\end{equation}
In the last step we have used that $D_{pp}\propto p^2$ for the turbulent
reacceleration (equation \ref{eqn:Dpp}). Hence the acceleration time is independent of momentum.}
%\begin{equation}
%\tau_{\rm D}= \frac{p^{2}}{4 D_{pp}}
%\end{equation}
%which is momentum independent.
This should be compared against the lifetime of turbulence, $\tau_{\rm cl}$, and the cooling time $\tau_{\rm cool}$. In Table~\ref{tab:timescales}, we show both the thermal quantities and the
timescales for CR cooling and (re)acceleration for three different
spatial regions of the RH. The densities \citep{1992A&A...259L..31B} and temperatures \citep{2009ApJ...696.1886B,2001A&A...365L..67A} are derived from X-ray observations. To calculate synchrotron cooling times, we use $B$-fields derived from Faraday rotation measurements \citep{bonafede10}. To calculate the acceleration time, we need to assume an outer scale. We assume an injection scale $k_L=2\upi/\lambda_L$, where $\lambda_L = 100$ kpc, which were assumptions adopted in previous work \citep{2006MNRAS.366.1437S,brunetti07, brunetti11}. This length scale corresponds to an eddy turnover time on the outer scale of $\varv_L k_L^{-1} \sim 1.2$~Gyr if $\varv_L \sim 500 {\rm km \, s^{-1}}$, as is characteristic of a merger. Note that hydrodynamical simulations of clusters have sometimes found larger $\lambda_L$, in some cases comparable to the size of the cluster \citep{vazza11,miniati15}. This choice is degenerate with $X_{\rm tu}$; in \S\ref{sect:self-limiting} we argue that $k_{\rm A}$ is a more appropriate choice, but also find that when the decay time is appropriately scaled, we are relatively insensitive to the choice of $k_L$. We present $\tau_{\rm D}$ for 3 different models, which we discuss in the next section. The reacceleration timescale
$\tau_{\rm D}$ is similar between our three models, where the
difference comes from turbulent profile parametrized with
$\alpha_\rmn{tu}$. This implies that even small differences in the
turbulent profile could impact the seed CRs significantly. Finally, we adopt an duration of acceleration $\tau_{\rm cl} = 650$ Myr, in line with previous assumptions \citep{brunetti07}, roughly corresponding to the turbulent decay time. 
\begin{table}
  \caption{Thermal quantities and timescales for different spatial
    regions in a Coma like cluster.}
\begin{tabular}{l c  c c}
\hline
\hline
& & spatial regions & \\
 & $0.1\,\RH^{(2)}$ & $0.3\,\RH^{(2)}$ & $\RH^{(2)}$   \\
\hline
thermal quantities$^{(1)}$ & & & \\
$\rho$ [$10^{-27}$ g cm$^{-3}$] & 3.0 & 1.6 & 0.15 \\
$T$ [$10^{8}$ K] & 1.4 & 1.0 & 0.58 \\
\hline
timescales$^{(1)}$ & & & \\
$\tau_\rmn{D}$(\Mprimary)$^{(3)}$ [Gyr] & 0.45 & 0.44 & 0.39 \\
$\tau_\rmn{D}$(\Mstream)$^{(3)}$  [Gyr] & 0.50 & 0.47 & 0.34 \\
$\tau_\rmn{D}$(\Mflatturb)$^{(3)}$  [Gyr] & 0.69 & 0.56 & 0.27 \\
$\tau_\rmn{IC/sync}(P=10^4\,m_{\rmn{e}}c)^{(4)}$ [Gyr] & 0.11 & 0.15 & 0.22 \\
$\tau_\rmn{had}(P=100\,m_{\rmn{p}}c)^{(5)}$ [Gyr] & 2.4 & 4.5 & 47 \\
$\tau_\rmn{Coul}(P=\,m_{\rmn{e}}c)^{(6)}$  [Gyr] & 0.0092 & 0.017 & 0.17 \\
\hline
\end{tabular}
\begin{quote}
 Notes: \\ 
 (1) Median quantities from our simulated post-merging cluster g72a derived 
 during last 300~Myrs in time. \\
 (2) Radius of the giant radio halo in Coma where $\RH\approx0.6\,R_{200}$.\\
 (3) Fermi-II reacceleration for both electrons and protons at all energies.\\
 (4) Inverse Compton and synchrotron cooling for electrons.\\
 (5) Catastrophic losses for protons.\\
 (6) Coulomb cooling for electrons (protons factor $m_{\rmn{e}}/m_{\rmn{p}}$ smaller).\\
 \label{tab:timescales}
  \end{quote}
\end{table}
